A Reduced Order Model for Permeability Fields Using Singular Value Decomposition (SVD)

### 2018-04-13 Papers/AppliedMaths2018/Stochastic_Upscaling_UK_AMM_2018.tex
updates

\section{Mathematical Model}\label{section:mathematical_model}

\citet{Gomes_2017} provides details on the governing equations and numerical formulation which forms the mathematical model for the multi-flow flow in this work. Albeit, it suffices to introduce only the representation of Darcy's law for immiscible multi-phase flow for the following discussion. This is written in the form:
\begin{equation}\label{e:darcy_eqn}
  \mathbf{q}_{\alpha} =
  -\frac{\mathcal{K}_{{r}_\alpha}\mathbf{K}}{\mu_{\alpha}}\left(
  \nabla p_{\alpha} - {\mathbf{s}_{u}}_{\alpha} \right),
\end{equation}
where $\alpha$ designates a phase and $\mathbf{q}_{\alpha}$ is the $\alpha$-th phase Darcy velocity. $\mu_{\alpha}$, $p_{\alpha}$, $\rho_{\alpha}$, and $\mathbf{s}_{{u}_\alpha}$ are the phase dynamic viscosity, pressure, density and source term, which may include gravity and/or capillarity, respectively. $\mathcal{K}_{{r}_\alpha}\left(S_{\alpha}\right)$ is the phase relative permeability, which is a function of the phase saturation $S_{\alpha}\left(\mathbf{r},t\right)$, which in turn is a function of the vector position $\mathbf{r}$, and time, $t$. Finally, $\mathbf{K}\left(\mathbf{r}\right)$, the focus of this work, is the absolute permeability tensor of the porous medium and it is a function of the vector position $\mathbf{r}$.

>>>>

\section{Multi-Fluid Flow Model}\label{section:mathematical_model}

In this work, a novel control volume finite element method (CVFEM) formulation is used to discretise and solve the set of multi-fluid flow equations. Continuity equations are embedded into the pressure equation to enforce mass conservation and the exact force balance (extended Darcy equations). The numerical formulation employs an implicit algorithm with respect to time which is less restrictive than the traditional industry-standard implicit-pressure-explicit-saturation (IMPES) scheme \cite{aziz_1986,geiger_2004}.

The formulation used in this work uses a dual consistent CVFEM representation which is embedded in novel families of triangular and tetrahedral finite element-pairs, \PN[n]{m} and \PNDG[n]{m}. In this element-pairs velocity is represented by $n^{\rm th}$-order polynomials that are discontinuous across elements, while pressure is represented by $m^{\rm th}$-order polynomials that may be either continuous (\PN[n]{m}) or discontinuous (\PNDG[n]{m}) across elements. Most properties of \PNDG[n]{m} element-pairs are similar to those in the \PN[n]{m} element-pairs but allows a representation in which pressure, saturation and other solution variables (\eg temperature, concentrations etc) are discontinuous across finite element boundaries \cite{salinas_2018,adam_2016}.

\cite{Gomes_2017} provides details on the governing equations and numerical formulation which forms the mathematical model for the multi-flow flow in this work. Albeit, it suffices to introduce only the representation of Darcy's law for immiscible multi-phase flow for the following discussion. This is written in the form:
\begin{equation}\label{e:darcy_eqn}
  \mathbf{q}_{\alpha} =
  -\frac{\mathcal{K}_{{r}_\alpha}\mathbf{K}}{\mu_{\alpha}}\left(
  \nabla p_{\alpha} - {\mathbf{s}_{u}}_{\alpha} \right),
\end{equation}
where $\alpha$ designates a phase and $\mathbf{q}_{\alpha}$ is the $\alpha$-th phase Darcy velocity. $\mu_{\alpha}$, $p_{\alpha}$, $\rho_{\alpha}$, and $\mathbf{s}_{{u}_\alpha}$ are the phase dynamic viscosity, pressure, density and source term, which may include gravity and/or capillarity, respectively. $\mathcal{K}_{{r}_\alpha}\left(S_{\alpha}\right)$ is the phase relative permeability, which is a function of the phase saturation $S_{\alpha}\left(\mathbf{r},t\right)$, which in turn is a function of the vector position $\mathbf{r}$, and time, $t$. Finally, $\mathbf{K}\left(\mathbf{r}\right)$, the focus of this work, is the absolute permeability tensor of the porous medium and it is a function of the vector position $\mathbf{r}$.
